\pdfoutput=1
\documentclass[10pt,logo,copyright]{nvidiatechreport}
\linespread{1.15}

\input{packages}
% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{common.tex}
\newcommand{\FIGREF}[1]{Fig.~\ref{#1}} % Fig. or Figure
\definecolor{darkred}{rgb}{0.7, 0.0, 0.0}
\newcommand{\checkthis}[1]{\textcolor{darkred}{#1}}

\newcommand{\red}[1]{\textcolor{red}{#1}}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}} % Tick mark
\newcommand{\xmark}{\ding{55}} % Cross mark
\usepackage{wrapfig}
\renewcommand{\today}{}
\usepackage[nameinlink]{cleveref}
\crefname{equation}{Eq.}{Eqs.}
\crefname{figure}{Fig.}{Figs.}
\crefname{section}{Sec.}{Sec.}
\crefname{appendix}{App.}{App.}
\crefname{table}{Tab.}{Tabs.}
\crefname{algorithm}{Algo}{Algo}
\crefname{thm}{Thm}{Thm}
\Crefname{thm}{Thm}{Thm}
\crefname{prop}{Prop}{Prop}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage{makecell}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{hyperref} 
\usepackage{tikz}
\usetikzlibrary{trees} % Explicitly load the trees library
\usepackage[edges]{forest}
\usepackage[breakable]{tcolorbox}
\definecolor{nvidiaGreen}{HTML}{9dca63}
\newcommand{\crefnames}[3]{%
  \@for\next:=#1\do{%
    \expandafter\crefname\expandafter{\next}{#2}{#3}%
  }%
}
\usepackage{listings}
\lstdefinestyle{pythonstyle}{language=Python, basicstyle=\ttfamily\small, keywordstyle=\color{blue}, commentstyle=\color{gray}, stringstyle=\color{red}, showstringspaces=false, breaklines=true}
\newcommand{\sz}[1]{\textcolor[rgb]{0.12,0.55,1.0}{Shaokun-#1}}
\newcommand\ximing[1]{{\color{purple}[#1]$_{Ximing}$}}
\definecolor{midnightgreen}{rgb}{0.0, 0.29, 0.33}
\definecolor{deepgreen}{HTML}{0aa344}
\definecolor{deeppurple}{HTML}{7030a0}
\definecolor{deepblue}{HTML}{171d91}
\definecolor{brown}{HTML}{843c0c}
\definecolor{shadered}{HTML}{ffe5e5}
\definecolor{shadegreen}{HTML}{e5f7ed}
\definecolor{msftBlack}{RGB}{0,0,0}
\definecolor{lightred}{RGB}{255,163,163}
\definecolor{deepred}{RGB}{146,0,0}
\captionsetup[subfigure]{justification=centering}
\newcommand{\finding}[2]{
    % args:
    %   1: finding number
    %   2: finding text
    \vspace{-0.1cm}
    \begin{tcolorbox}[
        colback=nvidiaGreen,     % Soft ivory background
        colframe=gray!5!white,     % Deep navy frame color
        arc=10pt,                    % Less rounded corners for a modern look
        boxsep=7pt,                 % Inner margin between text and frame
        left=10pt,                  % Left margin within the box
        right=10pt,                 % Right margin within the box
        top=2pt,                    % Top margin within the box
        bottom=2pt,                 % Bottom margin within the box
        boxrule=1.0pt,              % Frame thickness
        % enhanced jigsaw removed
    ]
        \paragraph{\textbf{\textit{Finding #1:}}} #2
    \end{tcolorbox}
}
\newtcolorbox{boxL}{
    fontupper = \color{black},
    rounded corners,
    arc = 6pt,
    colframe = black!50, 
    boxrule = 0pt, 
    bottomrule = 4.5pt ,
    breakable,
}
\newcommand{\prorlagent}{\textsc{ProRL Agent }}
% \title{ProRL Agent: Decoupled Rollout Infrastructure for RL Training of Multi-Turn LLM Agents}
\title{ProRL Agent: Rollout-as-a-Service for RL Training of Multi-Turn LLM Agents}
\author{Hao Zhang$^*$, Mingjie Liu$^*$, Shaokun Zhang$^*$, Songyang Han, Jian Hu, Zhenghui Jin, Yuchi Zhang, Shizhe Diao, Ximing Lu, Binfeng Xu, Zhiding Yu, Jan Kautz, Yi Dong
}

\begin{abstract}
% \ximing{we might want to start with the importance then the difficulty with training multi-turn LLM agents.}
Multi-turn LLM agents are increasingly important for solving complex, interactive tasks, and reinforcement learning (RL) is a key ingredient for improving their long-horizon behavior.
However, RL training requires generating large numbers of sandboxed rollout trajectories, and existing infrastructures often couple rollout orchestration with the training loop, making systems hard to migrate and maintain.
Under the \textit{rollout-as-a-service} philosophy, we present \prorlagent, a scalable infrastructure that serves the full agentic rollout lifecycle through an API service.
\prorlagent also provides standardized and extensible sandbox environments that support diverse agentic tasks in rootless HPC settings.
We validate \prorlagent through RL training on software engineering, math, STEM, and coding tasks. \prorlagent is open-sourced at \href{https://github.com/NVIDIA-NeMo/ProRL-Agent-Server}{ProRL Agent} and integrated as part of NVIDIA NeMo Gym.
\end{abstract}

% \begin{abstract}
% We present \textsc{ProRL Agent}, a scalable rollout infrastructure for RL training of multi-turn LLM agents.
% Under the \textit{rollout-as-a-service} philosophy, \textsc{ProRL Agent} exposes the full agentic rollout lifecycle as a unified HTTP service, allowing trainers to submit task instances and retrieve completed trajectories without handling any rollout-side logic.
% It also provides standardized and extensible sandbox environments for supporting diverse agentic tasks in rootless HPC settings.
% We validate \textsc{ProRL Agent} through RL training on software engineering, math, STEM, and coding tasks. \textsc{ProRL Agent} is open-sourced at \href{https://github.com/NVIDIA-NeMo/ProRL-Agent-Server}{ProRL Agent} and integrated as part of NVIDIA NeMo Gym.
% \end{abstract}

\begin{document}

\maketitle
\abscontent

\input{introduction}
\input{relatedwork}
\input{method}
\input{experiment}

\section{Conclusion}

In this work, we introduce \textsc{ProRL Agent}, a open-source scalable rollout infrastructure for HPC-native multi-turn agent training. 
By separating the entire rollout lifecycle from policy training, \textsc{ProRL Agent} improves modularity, scalability, and deployability for agent RL. Experiments across software engineering, STEM, math, and code agents demonstrate effective end-to-end RL training, with strong performance gains across multiple model scales. We release \textsc{ProRL Agent} as open source and as part of NVIDIA NeMo Gym, and leave richer environments and improved cluster-scale robustness to future work.

\clearpage
\newpage
\bibliographystyle{plainnat}
\bibliography{reference}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn
\include{appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}


